Why do we Eval?\\[2mm]\Large A large-scale study of Eval usage in R

### 2020-12-16 analysis/usage.Rmd
More readable bins
Fix corpus size
changes ot usage

---
title: "Usage metrics"
author: "Pierre Donat-Bouillud"
output:
    html_document:
        gallery: false
        toc: true
        toc_depth: 3
        toc_float: true
        df_print: paged
date: "`r Sys.Date()`"
editor_options: 
  chunk_output_type: console
params:
  base_dir: /var/lib/R/project-evalR/run/package-evals-traced.10/
  calls_packages_path: summarized-packages.fst
  corpus_path: corpus.fst
  static_evals_path: package-evals-static.csv
---

@@

jan <- FALSE

@@

if (jan == FALSE) {
  source("inc/paths.R")
}
source("inc/latextags.R")
source("inc/setup.R")
theme_set(theme_minimal())
if (jan) {
  TAGS_DIR <- PLOT_DIR <- "."
}
create_tags(path(TAGS_DIR, "usage_metrics.tex"), prefix = "", default = TRUE)

base_dir <- params$base_dir
# Set the paths, and customize for Jan.
if (jan) {
  base_dir <- "/mnt/code/eval/run/package-evals-traced.9/"
}
eval_calls_path <- paste0(base_dir, params$calls_packages_path)
corpus_path <- paste0(base_dir, params$corpus_path)
eval_static_path <- paste0(base_dir, params$static_evals_path)
```

@@

```{r read,include=F}
read_fst(eval_calls_path) %>% as_tibble() -> E_raw
read_fst(corpus_path) %>% as_tibble() -> C_raw
read_csv(eval_static_path) %>% semi_join(C_raw, by = "package") -> P
```

@@

```{r select-columns,echo=F}
E_raw %>% select(
  run_package = package, #      The package which was run to trigger the eval
  ev_variant = eval_function, # Which eval variant was called
  duplicates = nb_ev_calls, #   Number of identical eval calls (weight)
  src_ref = eval_call_srcref, # Source ref for the eval call site
  file, #                       Script file name (name of this run)
  ev_package = eval_source, #   Package of the call site
  arg_size = expr_expression_length, # Size in bytes of eval argument string
  opcodes = direct_interpreter_eval #  Count of the AST nodes traversed by the
) -> E #                               interpreter during eval (non-recursive).

C_raw -> C
corpus_size <- nrow(P)
```

@@

The corpus has `r corpus_size` packages, and the data set has `r nrow(E)` 
distinct calls to the `eval` and its variants. There are 
`r length(unique(E$ev_package))` corpus packages with `eval` calls (note that
we may not trigger all `eval`s in a package due to spotty covereage).

@@

TBL <- overview_table(
  r("nb eval and variant calls", nb_calls),
  r("nb eval calls", eval_calls),
  r("nb evalq calls", evalq_calls),
  r("nb eval parent calls", eval.parent_calls),
  r("nb local calls", local_calls)
)
```

@@

```{r eval-calls-summary}

eval_calls_table <- E %>% 
  count(ev_package, wt=duplicates, sort = TRUE)

breaks <- c(0, 10, 100, 1000, 10000, 100000, 1000000, 10000000, 100000000)

group_tags <- cut(eval_calls_table$n,
                  breaks = breaks,
                  include.lowest = TRUE)

ev_summary <- summary(group_tags)

r("bin1", ev_summary[[1]])
r("bin2", ev_summary[[2]])
r("bin3", ev_summary[[3]])
r("bin4", ev_summary[[4]])
r("bin5", ev_summary[[5]])
r("bin6", ev_summary[[6]])
r("bin7", ev_summary[[7]])
r("bin8", ev_summary[[8]])
```

>>>>

---
title: "Usage metrics"
author: "Pierre Donat-Bouillud"
output:
    html_document:
        gallery: false
        toc: true
        toc_depth: 3
        toc_float: true
        df_print: paged
date: "`r Sys.Date()`"
editor_options: 
  chunk_output_type: console
params:
  base_dir: Data/
  calls_packages_path: summarized-packages.fst
  corpus_path: corpus.fst
  static_evals_path: package-evals-static.csv
---

@@

@@

source("inc/paths.R")
source("inc/latextags.R")
source("inc/setup.R")
theme_set(theme_minimal())
create_tags(path(TAGS_DIR, "usage_metrics.tex"), prefix = "", default = TRUE)

base_dir <- params$base_dir
eval_calls_path <- paste0(base_dir, params$calls_packages_path)
corpus_path <- paste0(base_dir, params$corpus_path)
eval_static_path <- paste0(base_dir, params$static_evals_path)
```

@@

```{r read, message=TRUE, include=F}
read_fst(eval_calls_path) %>% as_tibble() -> E_raw
read_fst(corpus_path) %>% as_tibble() -> C_raw
read_csv(eval_static_path) %>% semi_join(C_raw, by = "package") -> P
```

@@

```{r select-columns, message=TRUE, warning=FALSE, include=FALSE}
E_raw %>% select(
  run_package = package, #      The package which was run to trigger the eval
  ev_variant = eval_function, # Which eval variant was called
  duplicates = nb_ev_calls, #   Number of identical eval calls (weight)
  src_ref = eval_call_srcref, # Source ref for the eval call site
  file, #                       Script file name (name of this run)
  ev_package = eval_source, #   Package of the call site
  arg_size = expr_expression_length, # Size in bytes of eval argument string
  opcodes = direct_interpreter_eval #  Count of the AST nodes traversed by the
) -> E #                               interpreter during eval (non-recursive).

C_raw -> C
corpus_size <- nrow(P)

overview_table(
  r("Corpus",corpus_size)
) 
```

The size of our corpus is `corpus_size` packages.

@@

The data set has `r nrow(E)`distinct calls to the `eval` and its variants. 
There are `r length(unique(E$ev_package))` corpus packages with `eval` calls
(note that we may not trigger all `eval`s in a package due to spotty covereage).

@@

overview_table(
  r("Allcalls", nb_calls),
  r("Evals", eval_calls),
  r("Evalqs", evalq_calls),
  r("Eparents", eval.parent_calls),
  r("Locals", local_calls),
  r("Triggeredpkgs", length(unique(E$ev_package)))
) -> TBL
```

@@

# Ev-dist

We compute the frequency of evals by package. This shows that some packages
hardly use eval while other use it intensily. The numbers are "for all runs",
this means that a popular package may be triggered by more runs and thus
have higher frequencies. We will later control for this by looking at individual 
call sites.

```{r ev-dist, include=FALSE}
eval_calls_table <- E %>% 
  count(ev_package, wt=duplicates, sort = TRUE)
breaks <- c(0, 10^seq(1,8))
group_tags <- cut(eval_calls_table$n,
                  breaks = breaks,
                  include.lowest = TRUE)
ev_summary <- summary(group_tags)

overview_table(
r("bina", ev_summary[[1]]),
r("binb", ev_summary[[2]]),
r("binc", ev_summary[[3]]),
r("bind", ev_summary[[4]]),
r("bine", ev_summary[[5]]),
r("binf", ev_summary[[6]]),
r("bing", ev_summary[[7]]),
r("binh", ev_summary[[8]])) -> TBL
```

The distribution of eval calls, each bin represents a power of 10, the number in the bin is the count of packages.

### 2020-12-16 paper/main.tex
changes ot usage

\documentclass[USenglish,cleveref, autoref, thm-restate]{lipics-v2019}
\usepackage{listings,hyperref,multirow,paralist,xspace,url}
\input{macro}
\newcommand{\mypara}[1]{\medskip\noindent\emph{#1}\xspace}
\newcommand{\NOTE}[1]{{\it Note: #1}\xspace}

@@

\begin{document}\maketitle

@@

This section presents a high-level picture of the usage of \eval in R.

(CHECK IF TRUE) As none of the Kaggle programs invoke \eval, the
remainder of the paper focuses solely on the CRAN data. We will return
to Kaggle in our conclusion.

We recorded \NbEvalAndVariantCallsRnd calls to \eval and its variants (\NbEvalCallsRnd
for \eval, \NbEvalqCallsRnd for \c{evalq}, \NbEvalParentCallsRnd for \c{eval.parent} and \NbLocalCallsRnd for
\c{local}).

>>>>

\documentclass[USenglish,cleveref, autoref, thm-restate]{lipics-v2019}
\usepackage{listings,hyperref,multirow,paralist,xspace,url,wrapfig}
\input{macro}
\newcommand{\mypara}[1]{\medskip\noindent\emph{#1}\xspace}
\newcommand{\NOTE}[1]{{\it Note: #1}\xspace}

@@

\begin{document}
\maketitle

@@

To provide a picture of \eval usage in R, we recorded \Allcalls
invocations of \eval and its variants in the CRAN corpus. These calls
were obtained by XXX runs, with runs that did not trigger \eval
discarded. Out of the \Corpus CRAN packages with \eval call sites, we
observed calls in \Triggeredpkgs packages, the difference can be chalk
down to incomplete code coverage and occasional failures due to our
pipeline.

\begin{wrapfigure}{r}{5.2cm}
\begin{tabular}{|r@{\,}r@{\,}lr|r@{\,}r@{\,}lr|} \hline
\tt 0 &--& \tt 10      & \Bina  & \tt 1K &--&\tt 100K  & \Bine\\
\tt 11 &--& \tt 100    & \Binb  & \tt 100K &--&\tt 1M  & \Binf\\
\tt 101 &--& \tt 1K    & \Binc  & \tt 1M &--&\tt 10M   & \Bing\\
\tt 1K &--& \tt 10K    & \Bind  & \tt 10M &--& \tt 100M & \Binh\\\hline
\end{tabular}
\caption{Call frequency}\label{freq}
\end{wrapfigure}

Figure~\ref{freq} shows the number of \evals generated by each package
over all runs that invoked functions in that package. Out of the XXX
packages in the corpus, XXX call have a low frequency of \eval calls,
with fewer than 100 of them. On the other hand, XXX packages call eval
more than 1,000 times, with one package (???) calling it XXXX times.

(\EvalsRnd for \eval, \EvalqsRnd for \c{evalq}, \EparentsRnd for
\c{eval.parent} and \LocalsRnd for \c{local}).
